```{r echo=FALSE,  message=FALSE, warning=FALSE}
library(tidyverse)
library(gridExtra)
library(ggsignif)
library(ggpointdensity)
library(ggpubr)
library(ggrepel)
```

# Four subpopulations bootstrapping 100 times

Here we show the results of the bootstrapping analysis did for the four subpopulations.  
We want to test if the signal between the whole population and subpopulations  
is consistent in the genes.  
To identify genes where local variation may be generating strong BLS signal, 
we have measured BLS in subpopulations.  
We have defined subpopulations based on the geographical clustering.  
The sizes of the subpopulations ranged between 22 and 45 and the total number  
of samples used was 133.  
To mitigate the effect of a reduced sample size, we bootstrapped the analysis 100.  
While keeping the same subpopulations size, we  randomly shuffled the samples  
across these four sets 100 times, thus generation 400 'in silico' subpopulations.
Next, we measured the two summary statistics (ballermix (CLR), and NCD1),  
combined them, and assign values to each gene.  
For the following visualizations we averaged the  
y-axis: Summary statistics mean across the 400 in silico subpopulations.
x-axis: Summary statistics for the totality of the 240 samples.


RICHARD wants 100 values per gene. 
Each of this values is the average of the summary statistic 
calculated with the 4 pops per set-1 to set-100.

I need to:
-Go into a set.  
-Calculate the average for the pop1-4.  
-Save it inside the corresponding set' directory.  
-Them combine the 100 values per gene into a master table with genes as rows and 
100 averaged-values as columns.  
-Calculate the sd for this 100 values.  
-Plot





## CLR

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Get the results from the heavy lifting done in R apocrita
from_apocrita_400 <- read.table("results/Master_CLR_combined_400.txt", 
                                header = TRUE)
from_apocrita_400$candens <- as.factor(from_apocrita_400$candens) 

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Information for ggplot anotation
gene1 <- subset(from_apocrita_400, gene == "LOC105207562")
gene2 <- subset(from_apocrita_400, gene == "LOC105200045")
gene3 <- subset(from_apocrita_400, gene == "LOC105202344")
gene4 <- subset(from_apocrita_400, gene == "LOC105199771")

three_genes <- rbind(gene1, gene4)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

clr_plot_400 <- ggplot(from_apocrita_400 %>% arrange(candens), 
       aes(x = clr, y = mean_clr, colour = sd_clr)) +
  geom_point(size = 3, alpha = 0.6) +
  # Adding linear regression line
  geom_smooth(method = "lm", se = TRUE, color = "grey") +  
  # Add names to selected points
  geom_label_repel(data = three_genes, aes(label = gene), colour = "black",
                   size= 2, nudge_x = 3, nudge_y = 6, direction = "both",
                   box.padding = 0.3, max.overlaps = 50) +
  # Change colors
  scale_colour_viridis_c(option = "D", direction = 1, name = "sd CLR") + 
  #Change axis labels
  labs(y = "Bootstraped 400 subpops", x = "Whole population", 
       subtitle = "CLR") + 
  #guides(shape = FALSE) +
  theme_minimal()
clr_plot_400
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Save the plot to the current working directory
ggsave("results/clr_plot_400.png", plot = clr_plot_400)

```

## NCD

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Get the results from the heavy lifting done in R apocrita
from_apocrita_400 <- read.table("results/Master_ncd05_combined_400.txt", header = TRUE)
from_apocrita_400$candens <- as.factor(from_apocrita_400$candens) 

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Information for ggplot annotation
gene1 <- subset(from_apocrita_400, gene == "LOC105207562")
gene2 <- subset(from_apocrita_400, gene == "LOC105200045")
gene3 <- subset(from_apocrita_400, gene == "LOC105202344")
gene4 <- subset(from_apocrita_400, gene == "LOC105199771")

three_genes <- rbind(gene1, gene4)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

ncd_plot_400 <- ggplot(from_apocrita_400 %>% arrange(candens), 
       aes(x = ncd05, y = mean_ncd05, colour = sd_ncd05)) +
  geom_point(size = 3, alpha = 0.6) +
  # Adding linear regression line
  geom_smooth(method = "lm", se = TRUE, color = "grey") +  
  # Add names to selected points
  geom_label_repel(data = three_genes, aes(label = gene), colour = "black",
                   size= 2, nudge_x = -0.1, nudge_y = -0.009, direction = "both",
                   box.padding = 0.3, max.overlaps = 50) +
  # Change colors
  scale_colour_viridis_c(option = "C", direction = 1, name = "sd ncd") + 
  #Change axis labels
  labs(y = "Bootstraped 400 subpops", x = "Whole population", 
       subtitle = "NCD") +
  #guides(shape = FALSE) +
  theme_minimal()
ncd_plot_400

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

 ggplot(from_apocrita_400 %>% subset(gene != "LOC105199771") %>%arrange(candens), 
       aes(x = ncd05, y = mean_ncd05, colour = sd_ncd05)) +
  geom_point(size = 3, alpha = 0.6) +
  # Adding linear regression line
  geom_smooth(method = "lm", se = TRUE, color = "grey") +  
  # Add names to selected points
  geom_label_repel(data = gene1, aes(label = gene), colour = "black",
                   size= 2, nudge_x = -0.01, nudge_y = -0.001, direction = "both",
                   box.padding = 0.3, max.overlaps = 50) +
  # Change colors
  scale_colour_viridis_c(option = "C", direction = 1, name = "sd ncd") + 
  #Change axis labels
  labs(y = "Bootstraped 400 subpops", x = "Whole population", 
       subtitle = "NCD no heparan") +
  #guides(shape = FALSE) +
  theme_minimal()

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Save the plot to the current working directory
ggsave("results/ncd_plot_400.png", plot = ncd_plot_400)

```

## Density

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Get the results from the heavy lifting done in R apocrita
from_apocrita_400 <- read.table("results/Master_density_combined_400.txt", header = TRUE)
from_apocrita_400$candens <- as.factor(from_apocrita_400$candens) 

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Information for ggplot anotation
gene1 <- subset(from_apocrita_400, gene == "LOC105207562")
gene2 <- subset(from_apocrita_400, gene == "LOC105200045")
gene3 <- subset(from_apocrita_400, gene == "LOC105202344")
gene4 <- subset(from_apocrita_400, gene == "LOC105199771")

three_genes <- rbind(gene1, gene4)

```



```{r echo=FALSE, message=FALSE, warning=FALSE}

density_plot_400 <- ggplot(from_apocrita_400 %>% arrange(clr), 
       aes(x = density, y = mean_density, colour = sd_density, shape = candens)) +
  geom_point(size = 3, alpha = 0.6) +
  # Adding linear regression line
##  geom_smooth(method = "lm", se = TRUE, color = "grey") +  
  # Add names to selected points
  geom_label_repel(data = three_genes, aes(label = gene), colour = "black",
                   size= 2, nudge_x = -0.0005, nudge_y = 0.002,  direction = "both",
                   box.padding = 0.3, max.overlaps = 50) +
  # Change colors
  scale_colour_viridis_c(option = "C", direction = -1, name = "sd in subpops") + #, 
                        #  labels = c("No","Yes")) + 
  #Change axis labels
  labs(y = "Bootstraped 400 subpops", x = "Whole population", 
       subtitle = "Combined statistics by density") + 
  #guides(shape = FALSE) +
  geom_hline(yintercept = quantile(from_apocrita_400$mean_density, 0.01), 
             linetype = "dashed", size = 0.5) +
  geom_vline(xintercept = quantile(from_apocrita_400$density, 0.01),
             linetype = "dashed", size = 0.5) +
  theme_minimal()
density_plot_400

```


From the plot above:
The button right quadrant is very interesting as the have strong signal in both all- and bootstrap-subpopultaion samples.
The quadrant made by the top 1% lines is a subset of our original candidate genes.
The show the stronger signal on the bootstrap analysis.


## 4 Real populations vs reshuffleted samples




