```{r echo=FALSE,  message=FALSE, warning=FALSE}
library(tidyverse)
library(gridExtra)
library(ggsignif)
library(ggpointdensity)
library(ggpubr)
library(ggrepel)
library(MASS)

```


```{r echo=FALSE,  message=FALSE, warning=FALSE}

bias_and_all <- read.table('input/2022-09-21-Master_info_Genes.txt',
                             header = TRUE) 
bias_and_all$sig <- ifelse(bias_and_all$clr >= 13.12954 & 
                              bias_and_all$ncd05 <= 0.4734869 , "1", "0")
bias_and_all$immune <- as.factor(as.numeric(bias_and_all$immune))

```

# 1. Plots with geom_pointdensity()

## This is the raw data.

```{r echo=FALSE, message=FALSE, warning=FALSE}

dens_clr_ncd_1 <- ggplot(bias_and_all , 
                         mapping = aes(x = clr, y = ncd05)) +
  geom_pointdensity(size = 3, shape = 18) + scale_colour_viridis_c( option = "C") + 
  theme_minimal()
  
#dens_clr_ncd_1

dens_clr_ncd_2 <- ggplot(data = bias_and_all, aes(x=clr, y=ncd05)) + 
  stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="white", bins = 20) + 
  theme_minimal()

#grid.arrange(dens_clr_ncd_1, dens_clr_ncd_2)
```

### Now I will scale them with the SD.
#### The SD should be calculated without the outliers

```{r echo=TRUE, message=FALSE, warning=FALSE}
#Filter the main data frame

no_outliers <- subset(bias_and_all, clr < 30)
no_outliers <- subset(no_outliers, ncd05 > 0.2)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

dens_clr_ncd_5 <- ggplot(no_outliers, mapping = aes(x = clr, y = ncd05)) +
  geom_pointdensity(size = 3, shape = 18) + scale_colour_viridis_c( option = "C") + 
  theme_minimal()
  
#dens_clr_ncd_2
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
#Calculate SD

sd_clr_no_outlier <- sd(no_outliers$clr)
sd_ncd_no_outlier <- sd(no_outliers$ncd05)

```

### Point-density plot scaled by the not-outliers SD 

```{r echo=FALSE, message=FALSE, warning=FALSE}

dens_clr_ncd_3 <- ggplot(bias_and_all, mapping = aes(x = clr/sd_clr_no_outlier,
                                                     y = ncd05/sd_ncd_no_outlier)) +
  geom_pointdensity(size = 3, shape = 18) + 
  scale_colour_viridis_c( option = "C") + 
  theme_minimal()
  
#dens_clr_ncd_3

dens_clr_ncd_4 <- ggplot(data = bias_and_all, aes(x=clr/sd_clr_no_outlier,
                                                  y=ncd05/sd_ncd_no_outlier)) + 
  stat_density_2d(aes(fill = ..level..), 
                  geom = "polygon", colour="white", bins = 20) + 
  theme_minimal()

#grid.arrange(dens_clr_ncd_3, dens_clr_ncd_4)
```

### This is the two of them side by side

```{r echo=FALSE, message=FALSE, warning=FALSE}

#grid.arrange(dens_clr_ncd_1, dens_clr_ncd_2,dens_clr_ncd_3, dens_clr_ncd_4, ncol =2)

```



####
### No outliers on y-axis. Bandwidth modified

The way to control the colors in the geom_pointdensity() method
is by modifying "adjust". The number of this option is multiplied by the default 
bandwidth. I don't like the fact that I can not find which is the value for 
their default bandwidth and I don't know how to access the calculated density values.
https://www.rdocumentation.org/packages/ggpointdensity/versions/0.1.0

```{r echo=FALSE, message=FALSE, warning=FALSE}

dens1 <- ggplot(bias_and_all %>% subset(gene != "LOC105199771"), 
                         mapping = aes(x = clr, y = ncd05)) +
  geom_pointdensity(size = 3, shape = 18) + scale_colour_viridis_c( option = "C") + 
  theme_minimal()
  
dens2 <- ggplot(data = bias_and_all %>% subset(gene != "LOC105199771")
                         , aes(x=clr, y=ncd05)) + 
  stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="white", bins = 20) + 
  theme_minimal()

#grid.arrange(dens1, dens2)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}

dens3 <- ggplot(bias_and_all %>% subset(gene != "LOC105199771"),
                mapping = aes(x = clr/sd_clr_no_outlier,
                              y = ncd05/sd_ncd_no_outlier)) +
  geom_pointdensity(adjust = 100, size = 4, shape = 18) + 
  scale_colour_viridis_c(option = "C") + 
  theme_minimal()

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
dens4 <- ggplot(data = bias_and_all %>% subset(gene != "LOC105199771"),
                aes(x=clr/sd_clr_no_outlier,
                    y=ncd05/sd_ncd_no_outlier)) + 
  stat_density_2d(aes(fill = ..level..), 
                  geom = "polygon", colour="white", bins = 20) + 
  theme_minimal()

#grid.arrange(dens3, dens4)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

grid.arrange(dens1, dens2, dens3, dens4, ncol =2)
dens3

```



#####################

# 2. Using MASS to calculate density
I like the fact that it adds the density values to the data frame.
https://slowkow.com/notes/ggplot2-color-by-density/

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(MASS)
library(ggplot2)
library(viridis)


#> Loading required package: viridisLite
theme_set(theme_bw(base_size = 16))

# Get density of points in 2 dimensions.
# @param x A numeric vector.
# @param y A numeric vector.
# @param n Create a square n by n grid to compute density.
# @return The density within each square.
get_density <- function(x, y, ...) {
  dens <- MASS::kde2d(x, y, ...)
  ix <- findInterval(x, dens$x)
  iy <- findInterval(y, dens$y)
  ii <- cbind(ix, iy)
  return(dens$z[ii])
}

set.seed(1)

```

#### I am playing arround with the following values:

"n" = Number of grid points in each direction. "I think 500 is good"

"h" = Vector of bandwidths for x and y directions. 



```{r echo=FALSE, message=FALSE, warning=FALSE}
### Option 1
# bias_and_all$density <- get_density(bias_and_all$clr/sd_clr_no_outlier, 
#                                     bias_and_all$ncd05/sd_ncd_no_outlier, 
#                                     h = c(9, 9), n = 500)
# 
# ggplot(bias_and_all %>% subset(gene != "LOC105199771"))+
#   geom_point(aes(clr, ncd05, color = density)) +
#   scale_color_viridis() +
#   theme_minimal()
# 
# nrow(subset(bias_and_all, density < 0.005))
# nrow(subset(bias_and_all, density < 0.002))

```



```{r echo=FALSE, message=FALSE, warning=FALSE}

### Option 2
# bias_and_all$density <- get_density(bias_and_all$clr/sd_clr_no_outlier, 
#                                     bias_and_all$ncd05/sd_ncd_no_outlier, 
#                                     h = c(13, 13), n = 500)
# 
# ggplot(bias_and_all %>% subset(gene != "LOC105199771"))+
#   geom_point(aes(clr, ncd05, color = density)) +
#   scale_color_viridis() +
#   theme_minimal()
# 
# nrow(subset(bias_and_all, density < 0.005))
# nrow(subset(bias_and_all, density < 0.002))
```


I like this one

```{r echo=FALSE, message=FALSE, warning=FALSE}

bias_and_all$density <- get_density(bias_and_all$clr/sd_clr_no_outlier, 
                                    bias_and_all$ncd05/sd_ncd_no_outlier, 
                                    h = c(17, 17), n = 500)

ggplot(bias_and_all %>% subset(gene != "LOC105199771")) +
  geom_point(aes(clr, ncd05, color = density), size = 4, shape = 18) +
  scale_color_viridis() +
  #geom_curve(aes(x = 3, y = 0.432, xend = 39, yend = 0.495), curvature = 0.4) +
  #geom_curve(aes(x = 3, y = 0.451, xend = 25, yend = 0.495), curvature = 0.55) +
  theme_minimal()

#nrow(subset(bias_and_all, density < 0.00436))
#nrow(subset(bias_and_all, density < 0.0015))

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# 
# bias_and_all$density <- get_density(bias_and_all$clr/sd_clr_no_outlier, 
#                                     bias_and_all$ncd05/sd_ncd_no_outlier, 
#                                     h = c(15, 15), n = 500)
# 
# ggplot(bias_and_all %>% subset(gene != "LOC105199771"))+
#   geom_point(aes(clr, ncd05, color = density)) +
#   scale_color_viridis() +
#   theme_minimal()
# 
# nrow(subset(bias_and_all, density < 0.0050))
# nrow(subset(bias_and_all, density < 0.0025))

```


### Candidate genes
```{r echo=FALSE, message=FALSE, warning=FALSE}

bias_and_all$candens <- "0"

for(i in 1:nrow(bias_and_all)) {                     
  if(is.na(bias_and_all$clr[i])){
    next
  }
  if(bias_and_all$density[i] <= 0.0041){
    bias_and_all$candens[i]  <- "1"
  }
  if(bias_and_all$density[i] <= 0.0015){
    bias_and_all$candens[i]  <- "2"
  }
}

bias_and_all$candens <- as.factor(bias_and_all$candens)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
bias_and_all$sig <- ifelse(bias_and_all$clr >= 13.12954 & 
                              bias_and_all$ncd05 <= 0.4734869 , "1", "0")

```

### Using density contours to select the candidates

```{r echo=FALSE, message=FALSE, warning=FALSE}

bins550 <- ggplot(bias_and_all %>% subset(gene != "LOC105199771") %>%
                    arrange(desc(density)),
             aes(clr/sd_clr_no_outlier, ncd05/sd_ncd_no_outlier)) + 
  geom_point(aes(colour = density), alpha = 0.7, size = 5, shape = 18) +
  scale_colour_viridis_c(option = "D", direction = 1, name = "Density") +
  theme_minimal()

bins550 + geom_density_2d(bins = 550, alpha = 0.3, color = "black") +
  xlab("B0-MAF") + ylab("NCD1") + 
  theme_minimal()


```


```{r echo=FALSE, message=FALSE, warning=FALSE}

bins550 <- ggplot(bias_and_all %>% arrange(desc(density)),
                  aes(clr/sd_clr_no_outlier, ncd05/sd_ncd_no_outlier)) + 
  geom_point(aes(colour = density), alpha = 0.7, size =3) +
  scale_colour_viridis_c(option = "D", direction = 1, name = "Density") +
  theme_minimal()

bins550 + geom_density_2d(bins = 550, alpha = 0.3, color = "black") +
  xlab("B0-MAF") + ylab("NCD1") + 
  theme_minimal()

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

my_genes <- ggplot(bias_and_all %>% subset(gene != "LOC105199771"),
             aes(clr/sd_clr_no_outlier, ncd05/sd_ncd_no_outlier)) + 
  geom_point(aes(colour = candens), alpha = 0.7, size = 5, shape = 18) +
  scale_colour_viridis_d(option = "D", direction = -1, name = "Density") +
  theme_minimal()

my_genes + geom_density_2d(bins = 550, alpha = 0.3, color = "black") +
  xlab("B0-MAF") + ylab("NCD1") + 
  theme_minimal()

```
```{r echo=FALSE, message=FALSE, warning=FALSE}

my_genes <- ggplot(bias_and_all %>% arrange(desc(density)),
             aes(clr/sd_clr_no_outlier, ncd05/sd_ncd_no_outlier)) + 
  geom_point(aes(colour = candens), alpha = 0.7, size =5, shape = 18) +
  scale_colour_viridis_d(option = "D", direction = -1, name = "Density") +
  theme_minimal()

my_genes + geom_density_2d(bins = 550, alpha = 0.3, color = "black") +
  xlab("B0-MAF") + ylab("NCD1") + 
  theme_minimal()


```

```{r echo=FALSE, message=FALSE, warning=FALSE}

nrow(subset(bias_and_all, density < 0.0041))
nrow(subset(bias_and_all, density < 0.0015))

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Get the SNPS for these 83 
# Add 1Kb in each direction
# Check if other genes are highlighted around these peaks in
# ~/btx645/2022-06-27-combine_tables_clr_ncd/2022-09-02-Gene_identification

top_peaks <- subset(bias_and_all, density <= 0.0041) 
top_peaks <- subset(top_peaks, select = c(chr, snp, gene)) %>% arrange(chr)
top_peaks$begin <- top_peaks$snp - 1000
top_peaks$end <- top_peaks$snp + 1000
top_peaks <- subset(top_peaks, select = c(chr, begin, end, gene, snp)) %>% arrange(chr)

write.table(top_peaks, "tmp/2023_05_31_Top_peaks.txt", quote = FALSE,
            col.names = FALSE, row.names = FALSE, sep = "\t")

##These are the extra ones that I got from the ana;ysis in 
# ~/2022-06-27-combine_tables_clr_ncd/2022-09-02-Gene_identification
# LOC120359458 LOC120359440
# LOC105196627 LOC105196643
# Trnap-agg LOC105204753
# LOC105196752 LOC105196820
# LOC105194434 LOC105194453

#subset(bias_and_all, gene == "LOC120359440")
#subset(bias_and_all, gene == "LOC105196643")
#subset(bias_and_all, gene == "LOC105196820")
#subset(bias_and_all, gene == "LOC105194453")

```



```{r echo=FALSE, message=FALSE, warning=FALSE}
#
top20_genes <- subset(bias_and_all, density <= 0.0015) 

write.table(top20_genes$gene, "tmp/2023_04_17_Top20_genes.txt", quote = FALSE,
            col.names = FALSE, row.names = FALSE)
#
top100_genes <- subset(bias_and_all, density <= 0.0041) 

write.table(top100_genes$gene, "tmp/2023_04_17_Top100_genes.txt", quote = FALSE,
            col.names = FALSE, row.names = FALSE)

#
No_candidates <- subset(bias_and_all, density > 0.0041) 

write.table(No_candidates$gene, "tmp/2023_04_17_No_candidates.txt", quote = FALSE,
            col.names = FALSE, row.names = FALSE)



write.table(bias_and_all, "results/2023_04_17_Dens_canditates.txt", quote = FALSE,
            col.names = TRUE, row.names = FALSE)


ranked_by_dens <- arrange(bias_and_all, density)

write.table(ranked_by_dens$gene, "results/2023_04_21_ranked_by_dens.txt", quote = FALSE,
            col.names = TRUE, row.names = FALSE)
```







----------------------
```{r echo=FALSE, message=FALSE, warning=FALSE}

# subTL <- ggplot(bias_and_all %>% subset(gene != "LOC105199771") %>% 
#             subset(clr < 2.8*sd_clr_no_outlier & ncd05 > 69.7*sd_ncd_no_outlier),
#              aes(clr/sd_clr_no_outlier, ncd05/sd_ncd_no_outlier)) + 
#   geom_point(aes(colour = candens), alpha = 0.7, size =3) +
#   scale_colour_viridis_d(option = "D", direction = 1) +
#   guides(colour = FALSE) +
#   theme_minimal() +
#   theme(axis.title.x=element_blank())
# 
# subTL <- subTL +geom_density_2d(bins = 20)
# #subTL

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

# subTR <- ggplot(bias_and_all %>% subset(gene != "LOC105199771") %>% 
#              subset(clr >= 2.8*sd_clr_no_outlier & ncd05 > 69.7*sd_ncd_no_outlier),
#              aes(clr/sd_clr_no_outlier, ncd05/sd_ncd_no_outlier)) + 
#   geom_point(aes(colour = candens), alpha = 0.7, size =3) +
#   scale_colour_viridis_d(option = "D", direction = 1) +
#   guides(colour = FALSE) +
#   theme_minimal() +
#   theme(axis.title.x=element_blank())
# 
# subTR <- subTR +geom_density_2d(bins = 20)
#subTR
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
#### FIX the axis here

# subBR <- ggplot(bias_and_all %>% subset(gene != "LOC105199771") %>% 
#              subset(clr >= 2.8*sd_clr_no_outlier & ncd05 <= 69.7*sd_ncd_no_outlier),
#              aes(clr/sd_clr_no_outlier, ncd05/sd_ncd_no_outlier)) + 
#   geom_point(aes(colour = candens), alpha = 0.7, size =3) +
#   scale_colour_viridis_d(option = "D", direction = 1) +
#   guides(colour = FALSE) +
#   theme_minimal() +
#   #xlim(2.8*sd_clr_no_outlier, 16.5) +
#   #ylim(69.7*sd_ncd_no_outlier, 60) +
#   theme(axis.title.x=element_blank(), axis.title.y=element_blank())
# 
# subBR <- subBR + geom_density_2d(bins = 20)
# #subBR
# #subp1 +geom_density_2d_filled(alpha = 0.5, show.legend = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

# subBL <- ggplot(bias_and_all %>% subset(gene != "LOC105199771") %>% 
#                   subset(clr < 2.8*sd_clr_no_outlier & ncd05 <= 69.7*sd_ncd_no_outlier) ,
#              aes(clr/sd_clr_no_outlier, ncd05/sd_ncd_no_outlier)) + 
#   geom_point(aes(colour = candens), alpha = 0.7, size =3) +
#   scale_colour_viridis_d(option = "D", direction = 1) +
#   guides(colour = FALSE) +
#  # xlim(0, 2.8*sd_clr_no_outlier) + 
#   theme_minimal() +
#   theme(axis.title.x=element_blank())
# 
# subBL <- subBL +geom_density_2d(bins = 20)
#subBL

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

# grid.arrange(subTL, subTR, subBL, subBR, ncol = 2)

```







###