
```{r echo=FALSE,  message=FALSE, warning=FALSE}
library(tidyverse)
library(gridExtra)
library(ggsignif)
library(ggpointdensity)
library(ggpubr)
library(ggrepel)
```


#### Is there a correlation between meassuring summary statistics (clr, ncd) for the  
#### totality of the 240 samples and those meassured in 4 sub-populations? 
x-axis: Summary statistics calculated with the totality of the 240 samples.  
y-axis: the average of the summary statistics calculated for the 4 subpopulations.  


```{r echo=FALSE, message=FALSE, warning=FALSE}
##### CLR

pop1_genes <- read.table('input/2023-08-04-ranked_by_dens-POP-1.txt', 
                         header = TRUE)
pop1_genes <- pop1_genes[, c("gene", "clr")]
names(pop1_genes) <- c("gene", "clr_pop1")
wide_pop1 <- pivot_wider(pop1_genes, names_from = "gene", 
                         values_from = "clr_pop1")

##
pop2_genes <- read.table('input/2023-08-04-ranked_by_dens-POP-2.txt', 
                         header = TRUE)
pop2_genes <- pop2_genes[, c("gene", "clr")]
names(pop2_genes) <- c("gene", "clr_pop2")
wide_pop2 <- pivot_wider(pop2_genes, names_from = "gene", 
                         values_from = "clr_pop2")

##
pop3_genes <- read.table('input/2023-08-04-ranked_by_dens-POP-3.txt', 
                         header = TRUE)
pop3_genes <- pop3_genes[, c("gene", "clr")]
names(pop3_genes) <- c("gene", "clr_pop3")
wide_pop3 <- pivot_wider(pop3_genes, names_from = "gene", 
                         values_from = "clr_pop3")

##
pop4_genes <- read.table('input/2023-08-04-ranked_by_dens-POP-4.txt', 
                         header = TRUE)
pop4_genes <- pop4_genes[, c("gene", "clr")]
names(pop4_genes) <- c("gene", "clr_pop4")
wide_pop4 <- pivot_wider(pop4_genes, names_from = "gene", 
                         values_from = "clr_pop4")

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#10864
merged_wide_clr <- bind_rows(wide_pop1, wide_pop2, wide_pop3, wide_pop4)

merged_wide_clr <- merged_wide_clr %>% select_if(~ !any(is.na(.)))
#6369
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

#pairs(t(merged_wide_clr))

#cor(t(merged_wide_clr))

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

bias_and_all <- read.table('input/2023_08_22_MASTER_cleaned_all_pops.txt',
                             header = TRUE)

common_loci <- names(merged_wide_clr)
means_pops_clrs <- colMeans(merged_wide_clr)
#combined_cls <- colMeans(merged_wide_clr)
#plot(means_pops_clrs,combined_clr,pch='.')
clrs = bias_and_all$clr
names(clrs) <- bias_and_all$gene

#plot(clrs[common_loci], means_pops_clrs, pch = ".")
#abline(30,-1)


clr_plot_data_to_combine <- data.frame(
  gene = common_loci,
  Clr = clrs[common_loci],
  Mean_Pops_Clr = means_pops_clrs
)

clr_plot_data <- merge(clr_plot_data_to_combine, bias_and_all, by= "gene")
clr_plot_data$candens <- as.factor(clr_plot_data$candens)

#locus = rep(names(merged_wide_clr),each=4)
#population = rep(paste("pop",1:4),dim(merged_wide_clr)[2])
#score <- as.vector(as.matrix(merged_wide_clr))
#longdat <-data.frame(locus,population,score)
#mod1 <- lm(score ~ 0+locus, data = longdat)

#head(coef(mod1))
#combined_clr <- coef(mod1)

write.table(clr_plot_data, "results/real_4_pops_clr.txt", 
            col.names = TRUE, quote = FALSE)


```


```{r echo=FALSE, message=FALSE, warning=FALSE}

clr_plot_average <- ggplot(clr_plot_data %>% arrange(candens), 
       aes(x = Clr, y = Mean_Pops_Clr, colour = candens)) +
  geom_point(size = 4, alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "grey") +  # Adding linear regression line
  #geom_label_repel(data = strong_genes, aes(label = gene), colour = "black",
   #                size= 2, nudge_x = 3, nudge_y = 6, direction = "both",
    #               box.padding = 0.3, max.overlaps = 50) +
  scale_colour_viridis_d(option = "D", 
                         direction = -1,
                        name = "Candidate",
                        labels = c("No","Yes")) + 
  labs(y = "Mean 4-pops") + 
  #guides(shape = FALSE) +
  theme_minimal()

```



```{r echo=FALSE, message=FALSE, warning=FALSE}

standardized_data <- scale(merged_wide_clr)

# Perform PCA
pca_result <- prcomp(standardized_data, center = TRUE, scale. = TRUE)

# Create a data frame with PCA scores
pca_scores <- as.data.frame(pca_result$x)

```



```{r echo=FALSE, message=FALSE, warning=FALSE}
# Explained variance 

explained_variance <- (pca_result$sdev^2) / sum(pca_result$sdev^2)


```


```{r echo=FALSE, message=FALSE, warning=FALSE}

pca_scores$row_number <- seq_len(nrow(pca_scores))

# Create a scatter plot with colored dots based on row number
pca_ballermix <- ggplot(pca_scores, aes(x = PC1, y = PC2, color = as.factor(row_number))) +
  scale_colour_viridis_d(option = "D", direction = 1, name = "SubPop") + 
  geom_point(size = 4) + labs(subtitle = "ballermix") +
  theme_minimal()

```




```{r echo=FALSE, message=FALSE, warning=FALSE}
##### NCD

pop1_genes <- read.table('input/2023-08-04-ranked_by_dens-POP-1.txt', 
                         header = TRUE)
pop1_genes <- pop1_genes[, c("gene", "ncd05")]
names(pop1_genes) <- c("gene", "ncd05_pop1")
wide_pop1 <- pivot_wider(pop1_genes, names_from = "gene", 
                         values_from = "ncd05_pop1")

##
pop2_genes <- read.table('input/2023-08-04-ranked_by_dens-POP-2.txt', 
                         header = TRUE)
pop2_genes <- pop2_genes[, c("gene", "ncd05")]
names(pop2_genes) <- c("gene", "ncd05_pop2")
wide_pop2 <- pivot_wider(pop2_genes, names_from = "gene", 
                         values_from = "ncd05_pop2")

##
pop3_genes <- read.table('input/2023-08-04-ranked_by_dens-POP-3.txt', 
                         header = TRUE)
pop3_genes <- pop3_genes[, c("gene", "ncd05")]
names(pop3_genes) <- c("gene", "ncd05_pop3")
wide_pop3 <- pivot_wider(pop3_genes, names_from = "gene", 
                         values_from = "ncd05_pop3")

##
pop4_genes <- read.table('input/2023-08-04-ranked_by_dens-POP-4.txt', 
                         header = TRUE)
pop4_genes <- pop4_genes[, c("gene", "ncd05")]
names(pop4_genes) <- c("gene", "ncd05_pop4")
wide_pop4 <- pivot_wider(pop4_genes, names_from = "gene", 
                         values_from = "ncd05_pop4")

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#10864
merged_wide_ncd05 <- bind_rows(wide_pop1, wide_pop2, wide_pop3, wide_pop4)

merged_wide_ncd05 <- merged_wide_ncd05 %>% select_if(~ !any(is.na(.)))
#6369
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

#pairs(t(merged_wide_ncd05))

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Compare the mean value fcalculated for each population agains the values 
# calculated for the totality of the samples

common_loci <- names(merged_wide_ncd05)
means_pops_ncd <- colMeans(merged_wide_ncd05)
ncd = bias_and_all$ncd05
names(ncd) <- bias_and_all$gene

ncd_plot_data_to_combine <- data.frame(
  gene = common_loci,
  ncd = ncd[common_loci],
  mean_pops_ncd = means_pops_ncd
)

ncd_plot_data <- merge(ncd_plot_data_to_combine, bias_and_all, by= "gene")
ncd_plot_data$candens <- as.factor(ncd_plot_data$candens)

write.table(ncd_plot_data, "results/real_4_pops_ncd.txt", 
            col.names = TRUE, quote = FALSE)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}

ncd05_plot_average <-ggplot(ncd_plot_data %>% arrange(candens), 
       aes(x = ncd05, y = mean_pops_ncd, colour = candens)) +
  geom_point(size = 4, alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "grey") +  # Adding linear regression line
  #geom_label_repel(data = strong_genes, aes(label = gene), colour = "black",
   #                size= 2, nudge_x = 3, nudge_y = 6, direction = "both",
    #               box.padding = 0.3, max.overlaps = 50) +
  scale_colour_viridis_d(option = "C", 
                         direction = -1, 
                         name = "Candidate",
                        labels = c("No","Yes")) + 
  labs(y = "Mean 4-pops") + 
  #guides(shape = FALSE) +
  theme_minimal()

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

standardized_data <- scale(merged_wide_ncd05)

# Perform PCA
pca_result <- prcomp(standardized_data, center = TRUE, scale. = TRUE)

# Create a data frame with PCA scores
pca_scores <- as.data.frame(pca_result$x)
#colnames(pca_scores) <- c("PC1", "PC2")  # Rename columns
```



```{r echo=FALSE, message=FALSE, warning=FALSE}
#Explained variance 

explained_variance <- (pca_result$sdev^2) / sum(pca_result$sdev^2)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

pca_scores$row_number <- seq_len(nrow(pca_scores))

# Create a scatter plot with colored dots based on row number
pca_ncd <- ggplot(pca_scores, aes(x = PC1, y = PC2, color = as.factor(row_number))) +
  scale_colour_viridis_d(option = "D", direction = 1, name = "SubPop") + 
  geom_point(size = 4) + labs(subtitle = "ncd") +
  theme_minimal()

```






```{r echo=FALSE, message=FALSE, warning=FALSE}
### Density-based (combined ncd and clr)

pop1_genes <- read.table('input/2023-08-04-ranked_by_dens-POP-1.txt', 
                         header = TRUE)
pop1_genes <- pop1_genes[, c("gene", "density")]
names(pop1_genes) <- c("gene", "density_pop1")
wide_pop1 <- pivot_wider(pop1_genes, names_from = "gene", 
                         values_from = "density_pop1")

##
pop2_genes <- read.table('input/2023-08-04-ranked_by_dens-POP-2.txt', 
                         header = TRUE)
pop2_genes <- pop2_genes[, c("gene", "density")]
names(pop2_genes) <- c("gene", "density_pop2")
wide_pop2 <- pivot_wider(pop2_genes, names_from = "gene", 
                         values_from = "density_pop2")

##
pop3_genes <- read.table('input/2023-08-04-ranked_by_dens-POP-3.txt', 
                         header = TRUE)
pop3_genes <- pop3_genes[, c("gene", "density")]
names(pop3_genes) <- c("gene", "density_pop3")
wide_pop3 <- pivot_wider(pop3_genes, names_from = "gene", 
                         values_from = "density_pop3")

##
pop4_genes <- read.table('input/2023-08-04-ranked_by_dens-POP-4.txt', 
                         header = TRUE)
pop4_genes <- pop4_genes[, c("gene", "density")]
names(pop4_genes) <- c("gene", "density_pop4")
wide_pop4 <- pivot_wider(pop4_genes, names_from = "gene", 
                         values_from = "density_pop4")

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#10864
merged_wide_density <- bind_rows(wide_pop1, wide_pop2, wide_pop3, wide_pop4)

merged_wide_density <- merged_wide_density %>% select_if(~ !any(is.na(.)))
#6369
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

#pairs(exp(t(merged_wide_density)))

#pairs(t(merged_wide_density))

#cor(t(merged_wide_density))
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Compare the mean value calculated for each population against the values 
# calculated for the totality of the samples

common_loci <- names(merged_wide_density)
means_pops_density <- colMeans(merged_wide_density)
density = bias_and_all$density
names(density) <- bias_and_all$gene

dens_plot_data <- data.frame(
  gene = common_loci,
  density_real = density[common_loci],
  means_pops_density = means_pops_density
)

dens_plot_data <- merge(dens_plot_data, bias_and_all, by= "gene")
dens_plot_data$candens <- as.factor(dens_plot_data$candens)

write.table(dens_plot_data, "results/real_4_pops_density.txt", 
            col.names = TRUE, quote = FALSE)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}

dens_plot_average <- ggplot(dens_plot_data %>% arrange(candens), 
       aes(x = density, y = means_pops_density, colour = candens)) +
  geom_point(size = 4, alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "grey") +  # Adding linear regression line
  #geom_label_repel(data = strong_genes, aes(label = gene), colour = "black",
   #                size= 2, nudge_x = 3, nudge_y = 6, direction = "both",
    #               box.padding = 0.3, max.overlaps = 50) +
  scale_colour_viridis_d(option = "A", 
                         direction = -1, 
                         name = "Candidate",
                        labels = c("No","Yes")) + 
  labs(y = "Mean 4-pops") + 
  #guides(shape = FALSE) +
  theme_minimal()

```



```{r echo=FALSE, message=FALSE, warning=FALSE}

standardized_data <- scale(merged_wide_density)

# Perform PCA
pca_result <- prcomp(standardized_data, center = TRUE, scale. = TRUE)

# Create a data frame with PCA scores
pca_scores <- as.data.frame(pca_result$x)

```




```{r echo=FALSE, message=FALSE, warning=FALSE}
#Explained variance 

explained_variance <- (pca_result$sdev^2) / sum(pca_result$sdev^2)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

pca_scores$row_number <- seq_len(nrow(pca_scores))

# Create a scatter plot with colored dots based on row number
pca_density <- ggplot(pca_scores, aes(x = PC1, y = PC2, color = as.factor(row_number))) +
  scale_colour_viridis_d(option = "D", direction = 1, name = "SubPop") + 
  geom_point(size = 4) + labs(subtitle = "Combined ncd+ballermix by density") +
  theme_minimal()

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

grid.arrange(clr_plot_average, ncd05_plot_average, dens_plot_average, ncol = 2)

```

##### These are PCAs with the goal to separate the populations.  
##### To check if the populations are independent.  
##### We could conclude that they are.  

```{r echo=FALSE, message=FALSE, warning=FALSE}

grid.arrange(pca_ballermix, pca_ncd, pca_density)

```



# Density-base selection

#####################

Using MASS to calculate density
https://slowkow.com/notes/ggplot2-color-by-density/

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(MASS)
library(ggplot2)
library(viridis)


#> Loading required package: viridisLite
theme_set(theme_bw(base_size = 16))

# Get density of points in 2 dimensions.
# @param x A numeric vector.
# @param y A numeric vector.
# @param n Create a square n by n grid to compute density.
# @return The density within each square.
get_density <- function(x, y, ...) {
  dens <- MASS::kde2d(x, y, ...)
  ix <- findInterval(x, dens$x)
  iy <- findInterval(y, dens$y)
  ii <- cbind(ix, iy)
  return(dens$z[ii])
}

set.seed(1)


means_for_density <- merge(
  ncd_plot_data_to_combine, 
  clr_plot_data_to_combine, 
  by = "gene")

sd_clr <- sd(means_for_density$Mean_Pops_Clr)
sd_ncd <- sd(means_for_density$mean_pops_ncd)


means_for_density$density <- get_density(means_for_density$Mean_Pops_Clr/sd_clr, 
                                    means_for_density$mean_pops_ncd/sd_ncd, 
                                    h = c(10, 10), n = 500)


```


```{r echo=FALSE, message=FALSE, warning=FALSE}
#%>% arrange(desc(density))
# ggplot(means_for_density  ,
#        aes(Mean_Pops_Clr, mean_pops_ncd)) + 
#   geom_point(aes(colour = density), alpha = 0.7, size = 5, shape =18) +
#   scale_colour_viridis_c(option = "D", direction = 1, name = "Density") +
#   #geom_density_2d(bins = 550, alpha = 0.3, color = "black") +
#   xlab("B0-MAF") + ylab("NCD") + 
#   theme_minimal()

```



### Candidate genes
##### The candiadate genes are those in the top 1% of the density distribution  
##### which was calculated with the mean of the summary statistics measured  
##### separately for the 4 subpopulations. 
##### These genes are shared by the 4 subpopulations and thus we belive the 
##### effect of environmental factors is diminished on them.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Select the top 1%

#quantile(means_for_density$density, 0.01)

means_for_density$candens <- "0"

for(i in 1:nrow(means_for_density)) {                     
  if(is.na(means_for_density$Mean_Pops_Clr[i])){
    next
  }
  if(means_for_density$density[i] <= quantile(means_for_density$density, 0.01)){
    means_for_density$candens[i]  <- "1"
  }
  # if(means_for_density$density[i] <= 0.0015){
  #   means_for_density$candens[i]  <- "2"
  # }
  # if(means_for_density$Mean_Pops_Clr[i] >= 16.11603 & means_for_density$ncd05[i] <= 0.4647506){
  #   means_for_density$candens[i]  <- "1"
  # }
}

means_for_density$candens <- as.factor(means_for_density$candens)

#Organize by Mean_Pops_Clr
means_for_density <- means_for_density[order(means_for_density$density, decreasing = FALSE), ]

#eliminate duplicates keeping the entry with the highest Mean_Pops_Clr per gene.
means_for_density <- means_for_density[!duplicated(means_for_density$gene), ]

#nrow(subset(means_for_density, candens == "1"))

```


```{r echo=TRUE, message=FALSE, warning=FALSE}

strong_genes <- subset(bias_and_all, PC1 >= 5) %>% arrange(desc(PC1))

write.table(strong_genes, file = "stong_genes_pca.txt", quote = FALSE,
            row.names = FALSE, col.names = FALSE)


ensemble_names <- read.table('pca_genes_ensembl.txt', 
                         header = TRUE, sep = "\t")

strong_genes <- merge(strong_genes, ensemble_names, by = "gene" )

#
gene1 <- subset(strong_genes, gene == "LOC105207562")
gene2 <- subset(strong_genes, gene == "LOC105200045")
gene3 <- subset(strong_genes, gene == "LOC105202344")

three_genes <- rbind(gene1, gene2, gene3)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#%>% subset(gene != "LOC105199771")
ggplot(means_for_density %>% 
  arrange(desc(density)), aes(Mean_Pops_Clr, mean_pops_ncd)) + 
  geom_point(aes(colour = density, shape = candens), alpha = 0.7, size = 5) +
  scale_colour_viridis_c(option = "D", direction = 1, name = "Density") +
  geom_label_repel(data = three_genes, aes(label = name), colour = "black",
                   size = 2, nudge_x = 2, nudge_y = 0.45, direction = "both",
                   box.padding = 0.3, max.overlaps = 50) +
  geom_density_2d(bins = 80, alpha = 0.3, color = "black") +
  labs(shape = "Candidates") + xlab("B0-MAF") + ylab("NCD1") + 
  scale_shape_manual(values = c(20, 18),
                     labels = c("No-BLS", "BLS")) +
  theme_minimal()

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#This are genes identified by:
# 1. Measuring CLR and NCD per sub-population
# 2. Calculate the means for these values 
# 3. Use the density "trick" to select those loci with lower density (the outliers)
nrow(subset(means_for_density, candens == "1"))

average_4_subpop_candidates <- subset(means_for_density, candens == "1") %>% arrange(density)
average_4_subpop_candidates$gene

write.table(average_4_subpop_candidates$gene,
            "tmp/average_4_subpop_cand_gene.txt",
            quote = FALSE, 
            row.names = FALSE,
            col.names = FALSE
            )

       
```