---
output:
  html_document: default
  pdf_document: default
---
```{r echo=FALSE,  message=FALSE, warning=FALSE}

library(tidyverse)
library(gridExtra)
library(ggsignif)
library(ggpointdensity)
library(ggpubr)
library(ggrepel)

```


```{r echo=FALSE,  message=FALSE, warning=FALSE}
#load("~/apocrita/2023-03-17-Caste_expression/common_tissues_caste_degs.RData")

bias_and_all <- read.table('../gitstuff/2023-Balancing_Selection/03-Candidate_genes_analysis/results/2023_08_22_MASTER_cleaned_all_pops.txt',
                             header = TRUE)
#bias_and_all$density <- bias_and_all$density*-1

#bias_and_all$immune <- as.factor(as.numeric(bias_and_all$immune))
bias_and_all$candens <- as.factor(as.numeric(bias_and_all$candens))


hyme_genes <- read.table('input/proteins_Hymenoptera_LOC.txt', header = FALSE)
bias_and_all$hymenop <- bias_and_all$gene %in% hyme_genes$V1
bias_and_all$hymenop <- as.factor(bias_and_all$hymenop*1)

```



# BUSCO hymenoptera

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(bias_and_all %>% subset(gene != "LOC105199771") %>% arrange(hymenop)) + 
  geom_jitter(aes(clr, ncd05, colour = hymenop, shape = candens), 
              alpha = 0.6, size = 6) +
  scale_colour_viridis_d(option = "E", direction = 1, name = "BUSCO", 
                         labels = c("No", "Yes")) +
  labs(shape = "Candidates") + xlab("B0-MAF") + ylab("NCD1") + 
  scale_shape_manual(values = c(20, 18), labels = c("No-BLS", "BLS")) +
  theme_minimal()

```


BUSCO genes in the BLS candidate genes

```{r echo = FALSE, message=FALSE, warning=FALSE}

busco_bls <- subset(bias_and_all, candens == "1") %>% subset(hymenop == "1")
busco_bls$gene
write.table(busco_bls$gene, 
            "../gitstuff/2023-Balaning_Selection/03-Candidate_genes_analysis/results/2023_07_22_BLS_BUSCO_genes.txt",
            quote = FALSE, col.names = FALSE, row.names = FALSE)
write.table(busco_bls, 
            "../gitstuff/2023-Balaning_Selection/03-Candidate_genes_analysis/results/2023_07_22_ALL_BUSCO.txt",
            quote = FALSE, col.names = TRUE, row.names = FALSE, sep = ",")

rm(busco_bls)

all_busco <- subset(bias_and_all, hymenop == "1")
write.table(all_busco$gene, 
            "../gitstuff/2023-Balaning_Selection/03-Candidate_genes_analysis/results/2023_07_22_ALL_BUSCO_genes.txt",
            quote = FALSE, col.names = FALSE, row.names = FALSE)

rm(all_busco)
```



Overrepresetnation?

```{r echo = FALSE, message=FALSE, warning=FALSE}

ggplot(bias_and_all, aes(x=hymenop, y=density, fill=hymenop)) +
  geom_violin(trim = TRUE) + geom_boxplot(width=0.1, fill="white") +
  labs(x = "Non BUSCO (7336) vs BUSCO (5220) genes", fill = "BUSCO") + 
  stat_compare_means(label.x = 1.3, label.y = -0.004)+
  scale_fill_discrete(labels = c("No", "Yes"))+
  theme_minimal() 
 
```


Fisher test




```{r echo = FALSE, message=FALSE, warning=FALSE}

fisher_busco <- data.frame(
 "NoBLS" = c(nrow(subset(bias_and_all, hymenop == "0") %>% subset(candens == "0")), 
            nrow(subset(bias_and_all, hymenop == "1") %>% subset(candens == "0"))),
 "BLS" = c(nrow(subset(bias_and_all, hymenop == "0") %>% subset(candens != "0")), 
            nrow(subset(bias_and_all, hymenop == "1") %>% subset(candens != "0"))),
  row.names = c("NoBUSCO", "BUSCO"),
  stringsAsFactors = FALSE
)

fisher_busco

fisher.test(fisher_busco)

```

```{r echo = FALSE, message=FALSE, warning=FALSE}
# Create a 2x2 contingency table with placeholder counts
data <- matrix(c(7249, 87, 5181, 39), nrow=2, byrow=TRUE) # Replace these numbers with your actual counts
rownames(data) <- c("NoBUSCO", "BUSCO")
colnames(data) <- c("NoBLS", "BLS")

# Convert the matrix to a data frame for ggplot
df <- as.data.frame(as.table(data))


p <- ggplot(df, aes(x=Var1, y=Freq, fill=Var2)) + 
  geom_bar(stat="identity", position="dodge") +
  labs(x="Group", y="Count", title="Feature Presence by Group") +
  scale_fill_manual(values=c("blue", "red")) +
  theme_minimal()

print(p)

test_result <- fisher.test(data)
print(test_result)

p_value <- test_result$p.value
mid_x <- mean(as.numeric(unique(df$Var1)))

p + annotate("text", x=mid_x, y=max(df$Freq)*0.8, label=paste("p-value:", round(p_value, 4)), vjust=1, hjust=1)


# Create the data frame
df <- as.data.frame(as.table(data))

# Bar plot with log scale
p_log <- ggplot(df, aes(x=Var1, y=Freq, fill=Var2)) +
  geom_bar(stat="identity", position="dodge") +
  scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),
                labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  labs(x="Presence of BUSCO", y="Count (Log Scale)", title="Distribution of BLS by BUSCO Presence") +
  scale_fill_manual(values=c("blue", "red")) +
  theme_minimal()

print(p_log)

install.packages("vcd")
library(vcd)

# Data setup
data <- matrix(c(7249, 87, 5181, 39), nrow=2, byrow=TRUE)
dimnames(data) <- list(BUSCO = c("NoBUSCO", "BUSCO"), BLS = c("NoBLS", "BLS"))

# Mosaic plot
mosaic(data, main="Mosaic Plot: Distribution of BLS by BUSCO Presence")

library(ggplot2)
library(scales)

# Create the data frame
df <- as.data.frame(as.table(data))

# Refined bar plot with log scale
p_refined <- ggplot(df, aes(x=Var1, y=Freq, fill=Var2)) +
  geom_bar(stat="identity", position="dodge") +
  geom_text(aes(label=Freq, y=1.1*Freq), position=position_dodge(width=0.9), vjust=-0.5, size=4) + # Add data labels
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x))) +
  labs(x="Presence of BUSCO", y="Count (Log Scale)", 
       title="Distribution of BLS by BUSCO Presence", 
       fill="Presence of BLS") +
  scale_fill_brewer(palette="Set1") + # Enhanced colors
  theme_minimal()

print(p_refined)


```

```{r echo = FALSE, message=FALSE, warning=FALSE}

library(ggplot2)
library(scales)
library(vcd)

# Your data matrix
data <- matrix(c(7249, 87, 5181, 39), nrow=2, byrow=TRUE)
rownames(data) <- c("NoBUSCO", "BUSCO")
colnames(data) <- c("NoBLS", "BLS")

# Convert the matrix to a data frame for ggplot
df <- as.data.frame(as.table(data))

# Compute the p-value using Fisher's exact test
fisher_test <- fisher.test(data)
p_value <- fisher_test$p.value


# Generate the mosaic plot
# Set the margins and then plot
par(mar=c(5.1, 4.1, 6, 2.1)) # Adjust these numbers as needed
# Define a color palette
color_palette <- c("red", "blue")

# Adjust the mosaic plot
assoc(data, 
      main=paste("Mosaic Plot\np-value:", sprintf("%.4f", p_value)),
      shade=TRUE,
      color=matrix(c("red", "blue"), nrow=2))




# Define shading matrix
shading_matrix <- matrix(c("lightblue", "blue", "lightpink", "red"), nrow=2, byrow=TRUE)

# Create the mosaic plot
mosaic(data, 
       main=paste("Mosaic Plot\np-value:", sprintf("%.4f", p_value)),
       gp=shading_fill(shading_matrix)
)


# Create a mosaic plot with the specified colors
mosaicplot(log(data), 
           main = paste("Mosaic Plot\np-value:", sprintf("%.4f", p_value)), 
           color = c("lightblue", "lightpink", "blue", "red"),
           las = 1, cex.main = 0.2)

```


```{r echo = FALSE, message=FALSE, warning=FALSE}
# Convert the matrix to a long format data frame for ggplot
df <- as.data.frame(as.table(data)) %>%
  group_by(Var1) %>%
  mutate(PosY = cumsum(Freq) - 0.5 * Freq)

# Create the mosaic plot
ggplot(df, aes(x=Var1, y=PosY, height=Freq, fill=Var2)) +
  geom_tile() +
  labs(
    title=paste("Mosaic Plot\np-value:", sprintf("%.4f", p_value)),
    x="Presence of BUSCO",
    y="Count",
    fill="Presence of BLS"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r echo = FALSE, message=FALSE, warning=FALSE}

# Refined bar plot with log scale
p_refined <- ggplot(df, aes(x=Var1, y=Freq, fill=Var2)) +
  geom_bar(stat="identity", position="dodge") +
  geom_text(aes(label=Freq, y=1.1*Freq), position=position_dodge(width=0.9), vjust=-0.5, size=4) +
  scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),
                labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  labs(title = "Distribution of BLS by BUSCO presence",
    x="Presence of BUSCO", y="Gene count (Log Scale)") +
  scale_fill_viridis_d(name = "BLS", labels = c("No", "Yes"), direction = -1) +
  annotate("text", x=1.5, y=max(df$Freq)/5,
           label=paste("p-value:", sprintf("%.4f", p_value)), hjust=1, vjust=0.5) +
  theme_minimal()

print(p_refined)

```



KS-test

```{r echo = FALSE, message=FALSE, warning=FALSE}
  
no_busco <- subset(bias_and_all, hymenop == "0")
yes_busco <- subset(bias_and_all, hymenop == "1")

ks.test(no_busco$clr, yes_busco$clr)

```
```{r echo = FALSE, message=FALSE, warning=FALSE}

#Perform the Wilcoxon rank-sum test
result <- wilcox.test(yes_busco$density, no_busco$density, alternative = "greater")
result

result <- wilcox.test(no_busco$density, yes_busco$density, alternative = "greater")
result
# Print the p-value
#print(result$p.value)

result <- wilcox.test(yes_busco$density, no_busco$density, alternative = "two.sided")
result

```
```{r echo=TRUE, message=FALSE, warning=FALSE}

ggplot() +
  geom_density(data = subset(bias_and_all, hymenop == "1"), 
               aes(x = density, fill = "BUSCO"), alpha = 0.5) +
  geom_density(data = subset(bias_and_all,hymenop == "0"), 
               aes(x = density, fill = "No-BUSCO"), alpha = 0.5) +
  xlab("BLS") + ylab("Density Estimate") + theme_minimal() +
  scale_fill_manual(values = c("blue", "red"), name = "Data Source", 
                    labels = c("BUSCO", "No-BUSCO"))

```


Only candidates 

```{r echo = FALSE, message=FALSE, warning=FALSE}

busco_vio_sig <- ggplot(bias_and_all %>% subset(candens != "0"), 
                        aes(x=hymenop, y=density, fill=hymenop)) +
  geom_violin(trim = TRUE) + geom_boxplot(width=0.1, fill="white") +
  labs(x = "Non BUSCO (59) vs BUSCO 24 genes", y = "CLR") + 
  stat_compare_means(label.x = 1.32, label.y = 0.0001)

busco_vio_sig + theme_minimal() 
#+ scale_fill_brewer(palette = "RdBu") + theme_minimal() 
```

KS-test only candidates

```{r echo = FALSE, message=FALSE, warning=FALSE}

no_busco <- subset(bias_and_all, candens != "0") %>% subset(hymenop == "0")
yes_busco <- subset(bias_and_all, candens != "0") %>%subset(hymenop == "1")

ks.test(no_busco$clr, yes_busco$clr)

```

#### BUSCO genes under BLS (25)

```{r echo = FALSE, message=FALSE, warning=FALSE}

candidate_25busco <- subset(bias_and_all, hymenop == "1") %>% subset(candens != "0")

write.table(candidate_25busco,
            file = "tmp/Complete_25_busco.csv", sep = ",", quote=FALSE,
            row.names = FALSE,  col.names = TRUE)

write.table(candidate_25busco$gene,
            file = "tmp/25-busco-candidates.csv", sep = ",", quote=FALSE,
            row.names = FALSE,  col.names = TRUE)

```

# BUSCO Insecta

```{r echo=FALSE,  message=FALSE, warning=FALSE}

insecta_genes <- read.table('input/proteins_Insecta_LOC.txt', header = FALSE)

bias_and_all$insecta <- bias_and_all$gene %in% insecta_genes$V1
bias_and_all$insecta <- as.factor(bias_and_all$insecta*1)

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(bias_and_all %>% subset(gene != "LOC105199771") %>% arrange(insecta)) + 
  geom_jitter(aes(clr, ncd05, colour = insecta, shape = candens), alpha = 0.6, size =5) +
  scale_colour_viridis_d(option = "D", direction = 1) +
  geom_curve(aes(x = 3, y = 0.451, xend = 30, yend = 0.495), curvature = 0.4) +
  theme_minimal()



```

Fisher test insecta

```{r echo = FALSE, message=FALSE, warning=FALSE}

fisher_insecta <- data.frame(
 "NoBLS" = c(nrow(subset(bias_and_all, insecta == "0") %>% subset(candens == "0")), 
            nrow(subset(bias_and_all, insecta == "1") %>% subset(candens == "0"))),
 "BLS" = c(nrow(subset(bias_and_all, insecta == "0") %>% subset(candens != "0")), 
            nrow(subset(bias_and_all, insecta == "1") %>% subset(candens != "0"))),
  row.names = c("NoBUSCO", "BUSCO"),
  stringsAsFactors = FALSE
)

fisher_insecta

fisher.test(fisher_insecta)

```

KS-test

```{r echo = FALSE, message=FALSE, warning=FALSE}
  
no_busco <- subset(bias_and_all, insecta == "0")
yes_busco <- subset(bias_and_all, insecta == "1")

ks.test(no_busco$clr, yes_busco$clr)

```

# BUSCO Diptera

```{r echo=FALSE,  message=FALSE, warning=FALSE}

diptera_genes <- read.table('input/proteins_Diptera_LOC.txt', header = FALSE)

bias_and_all$diptera <- bias_and_all$gene %in% diptera_genes$V1
bias_and_all$diptera <- as.factor(bias_and_all$diptera*1)

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(bias_and_all %>% subset(gene != "LOC105199771") %>% arrange(diptera)) + 
  geom_jitter(aes(clr, ncd05, colour = diptera, shape = candens), alpha = 0.6, size =5) +
  scale_colour_viridis_d(option = "D", direction = 1) +
  geom_curve(aes(x = 3, y = 0.451, xend = 30, yend = 0.495), curvature = 0.4) +
  theme_minimal()

```

Fisher test 

```{r echo = FALSE, message=FALSE, warning=FALSE}

fisher_diptera <- data.frame(
 "NoBLS" = c(nrow(subset(bias_and_all, diptera == "0") %>% subset(candens == "0")), 
            nrow(subset(bias_and_all, diptera == "1") %>% subset(candens == "0"))),
 "BLS" = c(nrow(subset(bias_and_all, diptera == "0") %>% subset(candens != "0")), 
            nrow(subset(bias_and_all, diptera == "1") %>% subset(candens != "0"))),
  row.names = c("NoBUSCO", "BUSCO"),
  stringsAsFactors = FALSE
)

fisher_diptera

fisher.test(fisher_diptera)

```

KS-test

```{r echo = FALSE, message=FALSE, warning=FALSE}
  
no_busco <- subset(bias_and_all, diptera == "0")
yes_busco <- subset(bias_and_all, diptera == "1")

ks.test(no_busco$clr, yes_busco$clr)

```


# BUSCO Lepidoptera

```{r echo=FALSE,  message=FALSE, warning=FALSE}

lepidop_genes <- read.table('input/proteins_Lepidoptera_LOC.txt', header = FALSE)

bias_and_all$lepidop <- bias_and_all$gene %in% lepidop_genes$V1
bias_and_all$lepidop <- as.factor(bias_and_all$lepidop*1)

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(bias_and_all %>% subset(gene != "LOC105199771") %>% arrange(lepidop)) + 
  geom_jitter(aes(clr, ncd05, colour = lepidop, shape = candens), alpha = 0.6, size =5) +
  scale_colour_viridis_d(option = "D", direction = 1) +
  geom_curve(aes(x = 3, y = 0.451, xend = 30, yend = 0.495), curvature = 0.4) +
  theme_minimal()

```

Fisher test 

```{r echo = FALSE, message=FALSE, warning=FALSE}

fisher_lepidop <- data.frame(
 "NoBLS" = c(nrow(subset(bias_and_all, lepidop == "0") %>% subset(candens == "0")), 
            nrow(subset(bias_and_all, lepidop == "1") %>% subset(candens == "0"))),
 "BLS" = c(nrow(subset(bias_and_all, lepidop == "0") %>% subset(candens != "0")), 
            nrow(subset(bias_and_all, lepidop == "1") %>% subset(candens != "0"))),
  row.names = c("NoBUSCO", "BUSCO"),
  stringsAsFactors = FALSE
)

fisher_lepidop

fisher.test(fisher_lepidop)

```

KS-test

```{r echo = FALSE, message=FALSE, warning=FALSE}
  
no_busco <- subset(bias_and_all, lepidop == "0")
yes_busco <- subset(bias_and_all, lepidop == "1")

ks.test(no_busco$clr, yes_busco$clr)

```


# BUSCO Osmia (solitary)

```{r echo=FALSE,  message=FALSE, warning=FALSE}

osmia_genes <- read.table('input/proteins_Osmia_LOC.txt', header = FALSE)

bias_and_all$osmia <- bias_and_all$gene %in% osmia_genes$V1
bias_and_all$osmia <- as.factor(bias_and_all$osmia*1)

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(bias_and_all %>% subset(gene != "LOC105199771") %>% arrange(osmia)) + 
  geom_jitter(aes(clr, ncd05, colour = osmia, shape = candens), alpha = 0.6, size =5) +
  scale_colour_viridis_d(option = "D", direction = 1) +
  geom_curve(aes(x = 3, y = 0.451, xend = 30, yend = 0.495), curvature = 0.4) +
  theme_minimal()

```

Fisher test 

```{r echo = FALSE, message=FALSE, warning=FALSE}

fisher_osmia <- data.frame(
 "NoBLS" = c(nrow(subset(bias_and_all, osmia == "0") %>% subset(candens == "0")), 
            nrow(subset(bias_and_all, osmia == "1") %>% subset(candens == "0"))),
 "BLS" = c(nrow(subset(bias_and_all, osmia == "0") %>% subset(candens != "0")), 
            nrow(subset(bias_and_all, osmia == "1") %>% subset(candens != "0"))),
  row.names = c("NoBUSCO", "BUSCO"),
  stringsAsFactors = FALSE
)

fisher_osmia

fisher.test(fisher_osmia)

```

KS-test

```{r echo = FALSE, message=FALSE, warning=FALSE}
  
no_busco <- subset(bias_and_all, lepidop == "0")
yes_busco <- subset(bias_and_all, lepidop == "1")

ks.test(no_busco$clr, yes_busco$clr)

```




Venn diagram

```{r echo=FALSE, message=FALSE, warning=FALSE}

hymeno_venn <- subset(bias_and_all, hymenop == "1")
insect_venn <- subset(bias_and_all, insecta == "1")
diptera_venn <- subset(bias_and_all, diptera == "1")
lepidop_venn <- subset(bias_and_all, lepidop == "1")
osmia_venn <- subset(bias_and_all, osmia == "1")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a Venn diagram 
library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(hymeno_venn$gene, insect_venn$gene, diptera_venn$gene, lepidop_venn$gene, osmia_venn$gene),
  category = c("Hymenoptera", "Insecta", "Diptera", "Lepidoptera", "Osmia"),
  fill = c("#2A8DA1", "#C0D9AE", "#B7BAB2", "#C0D9AE", "#C0D9AE"),
  height = 1000, width = 1000,
  filename = NULL,
  main = "BUSCO"
)

grid::grid.draw(venn_p1)
```


## Get the BUSCO genes exclusive to social hymenopteras

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Get the BUSCO genes exclusive to social insects

social_busco <- subset(bias_and_all, hymenop == "1") %>% subset(insecta == "0") %>% 
                  subset(diptera == "0") %>% subset(lepidop == "0") %>% subset(osmia == "0")

### Add a column on Bias_and_all with the social genes

social_busco <- as.data.frame(social_busco$gene)

bias_and_all$busco_social <- bias_and_all$gene %in% social_busco$`social_busco$gene`
bias_and_all$busco_social <- as.factor(bias_and_all$busco_social*1)

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(bias_and_all %>% subset(gene != "LOC105199771") %>% arrange(busco_social)) + 
  geom_jitter(aes(clr, ncd05, colour = busco_social, shape = candens), alpha = 0.6, size =5) +
  scale_colour_viridis_d(option = "D", direction = 1) +
  geom_curve(aes(x = 3, y = 0.451, xend = 30, yend = 0.495), curvature = 0.4) +
  theme_minimal()

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

#Get the shared BUSCOs between hymenoptera and any other category

bias_and_all$other_busco <- "0"

for(i in 1:nrow(bias_and_all)) {                     
  if(is.na(bias_and_all$clr[i])){
    next
  }
  if(bias_and_all$hymenop[i] == "1" & bias_and_all$insecta[i] == "1"){
    bias_and_all$other_busco[i]  <- "shared"
  }
  if(bias_and_all$hymenop[i] == "1" & bias_and_all$diptera[i] == "1"){
    bias_and_all$other_busco[i]  <- "shared"
  }
  if(bias_and_all$hymenop[i] == "1" & bias_and_all$lepidop[i] == "1"){
    bias_and_all$other_busco[i]  <- "shared"
  }
  if(bias_and_all$hymenop[i] == "1" & bias_and_all$osmia[i] == "1"){
    bias_and_all$other_busco[i]  <- "shared"
  }
}

bias_and_all$other_busco <- as.factor(bias_and_all$other_busco)


#Diffine thre categories for vizualisation. 
#Private for hymenoptera
#Shared with any other busco group
#No BUSCO gene


bias_and_all$busco_category <- "03-Other_BUSCO"

for(i in 1:nrow(bias_and_all)) {                     
  if(is.na(bias_and_all$clr[i])){
    next
  }
  if(bias_and_all$busco_social[i] == "1"){
    bias_and_all$busco_category[i]  <- "01-Private"
  }
  if(bias_and_all$hymenop[i] == "0" & bias_and_all$insecta[i] == "0" &
     bias_and_all$diptera[i] == "0" & bias_and_all$lepidop[i] == "0" & 
     bias_and_all$osmia[i] == "0"){
    bias_and_all$busco_category[i]  <- "04-No_BUSCO"
  }
  if(bias_and_all$other_busco[i] == "shared"){
    bias_and_all$busco_category[i]  <- "02-Shared"
  }
}

bias_and_all$busco_category <- as.factor(bias_and_all$busco_category)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

# ggplot(bias_and_all %>% subset(gene != "LOC105199771") %>% arrange(desc(busco_category))) + 
#   geom_jitter(aes(clr, ncd05, colour = busco_category, shape = candens), alpha = 0.8, size =5) +
#   scale_colour_viridis_d(option = "D", direction = -1) +
# #  geom_curve(aes(x = 3, y = 0.451, xend = 30, yend = 0.495), curvature = 0.4) +
#   theme_minimal()
# 

#write.table(subset(bias_and_all, hymenop == "1") %>% subset(candens != "0") %>%
 #             arrange(density),
  #          file = "tmp/24-busco-candidates.csv", sep = ",", quote=FALSE,
   #         row.names = FALSE,  col.names = TRUE)



# 
# ggplot(bias_and_all %>% subset(gene != "LOC105199771") %>% arrange(desc(busco_category))) + 
#   geom_jitter(aes(clr, ncd05, colour = density, shape = busco_category), alpha = 0.8, size =5) +
#   scale_colour_viridis_c(option = "D", direction = 1) +
#   geom_curve(aes(x = 3, y = 0.451, xend = 30, yend = 0.495), curvature = 0.4) +
#   theme_minimal()
# 
# 
# 
# ggplot(bias_and_all %>% subset(gene != "LOC105199771") %>% 
#                          subset(density < 0.006)%>%
#                          arrange(desc(busco_category))) + 
#   geom_jitter(aes(clr, ncd05, colour = density, shape = busco_category), alpha = 0.8, size =5) +
#   scale_colour_viridis_c(option = "D", direction = 1) +
#   #geom_curve(aes(x = 3, y = 0.451, xend = 30, yend = 0.495), curvature = 0.4) +
#   theme_minimal()
# 
# 
# nrow(subset(bias_and_all, busco_category == "01-Private") %>% subset(density < 0.006))

#subset(bias_and_all, busco_social == "1") %>% subset(candens != 0)
```

Fisher test

```{r echo = FALSE, message=FALSE, warning=FALSE}

fisher_lepidop <- data.frame(
 "NoBLS" = c(nrow(subset(bias_and_all, busco_social == "0") %>% subset(candens == "0")), 
            nrow(subset(bias_and_all, busco_social == "1") %>% subset(candens == "0"))),
 "BLS" = c(nrow(subset(bias_and_all, busco_social == "0") %>% subset(candens != "0")), 
            nrow(subset(bias_and_all, busco_social == "1") %>% subset(candens != "0"))),
  row.names = c("NoBUSCO", "BUSCO-private"),
  stringsAsFactors = FALSE
)

fisher_lepidop

fisher.test(fisher_lepidop)

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# #Print the BUSCO genes private to social insects which are candidates of BLS. 
# private_7_genes <- subset(bias_and_all, busco_social == "1") %>% subset(candens != 0)%>%
#                    arrange(density)
# 
# write.table(private_7_genes$gene, file = "tmp/7-private-busco-candidates.csv", 
#             sep = ",", 
#             quote=FALSE,
#             row.names = FALSE,
#             col.names = TRUE)
# 
# 
# private.2k <-subset(bias_and_all, busco_social == "1") %>%
#               arrange(density)
#   
# write.table(private.2k$gene, file = "tmp/2k-private-busco-candidates.csv", 
#             sep = ",", 
#             quote=FALSE,
#             row.names = FALSE,  
#             col.names = FALSE)
# 
# 
# write.table(subset(bias_and_all, candens != "0"), file = "tmp/BLS_genes_incomplete.csv", 
#             sep = ",", 
#             quote=FALSE,
#             row.names = FALSE,
#             col.names = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

# ggplot(bias_and_all %>% subset(gene != "LOC105199771") %>% arrange(density)) + 
#   geom_jitter(aes(clr, ncd05, colour = busco_category, shape = candens), alpha = 0.8, size =5) +
#   scale_colour_viridis_d(option = "D", direction = -1) +
#   geom_curve(aes(x = 3, y = 0.451, xend = 30, yend = 0.495), curvature = 0.4) +
#   theme_minimal()

```