```{r echo=FALSE,  message=FALSE, warning=FALSE}

library(tidyverse)
library(gridExtra)
library(ggsignif)
library(ggpointdensity)
library(ggpubr)
library(ggrepel)

```


```{r echo=FALSE,  message=FALSE, warning=FALSE}
#load("~/apocrita/2023-03-17-Caste_expression/common_tissues_caste_degs.RData")

bias_and_all <- read.table('results/2023_04_17_Dens_canditates.txt',
                             header = TRUE) 
bias_and_all <- subset(bias_and_all, chr != "CM031916.1")
bias_and_all$immune <- as.factor(as.numeric(bias_and_all$immune))
bias_and_all$candens <- as.factor(as.numeric(bias_and_all$candens))

hyme_genes <- read.table('input/proteins_Hymenoptera_LOC.txt', header = FALSE)
bias_and_all$hymenop <- bias_and_all$gene %in% hyme_genes$V1
bias_and_all$hymenop <- as.factor(bias_and_all$hymenop*1)

supergene_genes <- read.table('input/supergenes_LOC.txt', header = FALSE)
bias_and_all$super <- bias_and_all$gene %in% supergene_genes$V1
bias_and_all$super <- as.factor(bias_and_all$super*1) 

ase_genes <- read.table('input/consistent_AS_genes.txt', header = FALSE)
bias_and_all$ASE <- bias_and_all$gene %in% ase_genes$V1
bias_and_all$ASE <- as.factor(bias_and_all$ASE*1)

ors_genes <- read.table('input/LOCs_for_ORs.txt', header = FALSE)
bias_and_all$ors <- bias_and_all$gene %in% ors_genes$V1
bias_and_all$ors <- as.factor(bias_and_all$ors*1)

drosophila_genes <- read.table('tmp/inv_homologs_dm_bls.txt', header = FALSE)
bias_and_all$drosophila <- bias_and_all$gene %in% drosophila_genes$V1
bias_and_all$drosophila <- as.factor(bias_and_all$drosophila*1)

##
pop1_genes <- read.table('input/2023-08-04-POP-1-Candidate_Genes_top1pc.txt', header = FALSE)
bias_and_all$pop1 <- bias_and_all$gene %in% pop1_genes$V1
bias_and_all$pop1 <- as.factor(bias_and_all$pop1*1)

pop2_genes <- read.table('input/2023-08-04-POP2_Candidate_Genes.txt', header = FALSE)
bias_and_all$pop2 <- bias_and_all$gene %in% pop2_genes$V1
bias_and_all$pop2 <- as.factor(bias_and_all$pop2*1)

pop3_genes <- read.table('input/2023-08-04-POP-3_Candidate_Genes.txt', header = FALSE)
bias_and_all$pop3 <- bias_and_all$gene %in% pop3_genes$V1
bias_and_all$pop3 <- as.factor(bias_and_all$pop3*1)

pop4_genes <- read.table('input/2023-08-04-POP4_Candidate_Genes.txt', header = FALSE)
bias_and_all$pop4 <- bias_and_all$gene %in% pop4_genes$V1
bias_and_all$pop4 <- as.factor(bias_and_all$pop4*1)


```

```{r echo=FALSE,  message=FALSE, warning=FALSE}

candidate_immune <- subset(bias_and_all, immune == "1" & candens != "0")
candidate_busco <- subset(bias_and_all, hymenop == "1" & candens != "0")
candidate_super <- subset(bias_and_all, super == "1" & candens != "0")
candidate_ASE <- subset(bias_and_all, ASE == "1" & candens != "0")
candidate_ORs <- subset(bias_and_all, ors == "1" & candens != "0")
candidate_dros <- subset(bias_and_all, drosophila == "1" & candens != "0")

BLS_candidates <- subset(bias_and_all, candens != "0")

pop1_genes <- subset(bias_and_all, pop1 == "1")
pop2_genes <- subset(bias_and_all, pop2 == "1")
pop3_genes <- subset(bias_and_all, pop3 == "1")
pop4_genes <- subset(bias_and_all, pop4 == "1")

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
#top20
bias_and_all <- arrange(bias_and_all, density)
top20 <- head(bias_and_all, 20)
```

# All pops and samples 
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a Venn diagram 
library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(pop1_genes$gene, pop2_genes$gene, pop3_genes$gene, 
           pop4_genes$gene, BLS_candidates$gene),
  category = c("Pop1","Pop2", "Pop3", 
               "Pop4", "All-samples"),
  fill = c("#2A8DA1", "#C0D9AE", 
           "#B7BAB2", "#B45639", "#C0D9AE"),
  height = 1000, width = 1000,
  filename = NULL,
  main = "BLS"
)

grid::grid.draw(venn_p1)
```

# pop2

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a Venn diagram 
library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(BLS_candidates$gene, pop2_genes$gene),
  category = c("All", "Pop2"),
  fill = c("#B45639", "#2A8DA1"),
  height = 1000, width = 1000,
  filename = NULL,
  main = "BLS"
)

grid::grid.draw(venn_p1)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a Venn diagram 
library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(BLS_candidates$gene, pop3_genes$gene),
  category = c("All", "Pop3"),
  fill = c("#B45639", "#B7BAB2"),
  height = 1000, width = 1000,
  filename = NULL,
  main = "BLS"
)

grid::grid.draw(venn_p1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a Venn diagram 
library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(BLS_candidates$gene, pop4_genes$gene),
  category = c("All", "Pop4"),
  fill = c("#B45639", "#2A8DA1"),
  height = 1000, width = 1000,
  filename = NULL,
  main = "BLS"
)

grid::grid.draw(venn_p1)
```



```{r echo=FALSE, message=FALSE, warning=FALSE}

intersect(intersect(intersect(pop2_genes$gene, pop3_genes$gene), 
                    pop4_genes$gene), BLS_candidates$gene)

```
## Top 5% of population loci

```{r echo=FALSE, message=FALSE, warning=FALSE}
##
pop1_genes <- read.table('input/2023-08-04-POP-1-Candidate_Genes_top5pc.txt', 
                         header = FALSE)
bias_and_all$pop1 <- bias_and_all$gene %in% pop1_genes$V1
bias_and_all$pop1 <- as.factor(bias_and_all$pop1*1)

pop2_genes <- read.table('input/2023-08-04-POP2_Candidate_Genes_5pc.txt', 
                         header = FALSE)
bias_and_all$pop2 <- bias_and_all$gene %in% pop2_genes$V1
bias_and_all$pop2 <- as.factor(bias_and_all$pop2*1)

pop3_genes <- read.table('input/2023-08-04-POP-3_Candidate_Genes_top5pc.txt', 
                         header = FALSE)
bias_and_all$pop3 <- bias_and_all$gene %in% pop3_genes$V1
bias_and_all$pop3 <- as.factor(bias_and_all$pop3*1)

pop4_genes <- read.table('input/2023-08-04-POP4_Candidate_Genes_top5pc.txt',
                         header = FALSE)
bias_and_all$pop4 <- bias_and_all$gene %in% pop4_genes$V1
bias_and_all$pop4 <- as.factor(bias_and_all$pop4*1)

pop1_genes <- subset(bias_and_all, pop1 == "1")
pop2_genes <- subset(bias_and_all, pop2 == "1")
pop3_genes <- subset(bias_and_all, pop3 == "1")
pop4_genes <- subset(bias_and_all, pop4 == "1")

```

pop1-pop2

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a Venn diagram 
library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(pop1_genes$gene, pop2_genes$gene),
  category = c("Pop1","Pop2"),
  fill = c("#187E93", "#80ab61"),
  height = 1000, width = 1000,
  filename = NULL,
  main = "BLS"
)

grid::grid.draw(venn_p1)
```

pop1-pop4

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a Venn diagram 
library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(pop1_genes$gene, pop4_genes$gene),
  category = c("Pop1", "Pop4"),
  fill = c("#187E93", "#D7C16A"),
  height = 1000, width = 1000,
  filename = NULL,
  main = "BLS"
)

grid::grid.draw(venn_p1)
```

pop2-pop3

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a Venn diagram 
library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(pop2_genes$gene, pop3_genes$gene),
  category = c("Pop2", "Pop3"),
  fill = c("#80ab61", "#C0D9AE"),
  height = 1000, width = 1000,
  filename = NULL,
  main = "BLS"
)

grid::grid.draw(venn_p1)
```

pop2-pop4

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a Venn diagram 
library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(pop2_genes$gene, pop4_genes$gene),
  category = c("Pop2", "Pop4"),
  fill = c("#80ab61", "#D7C16A"),
  height = 1000, width = 1000,
  filename = NULL,
  main = "BLS"
)

grid::grid.draw(venn_p1)
```

pop3-pop4

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a Venn diagram 
library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(pop3_genes$gene, pop4_genes$gene),
  category = c("Pop3", "Pop4"),
  fill = c("#C0D9AE", "#D7C16A"),
  height = 1000, width = 1000,
  filename = NULL,
  main = "BLS"
)

grid::grid.draw(venn_p1)
```

# Only 4 populations

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Colours:
# All samples #7d55a8
# 1 #187e93
# 2 #4d9817
# 3 #ade187
# 4 #f1ec09


# Create a Venn diagram 
library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(pop1_genes$gene, 
           pop2_genes$gene, 
           pop3_genes$gene, 
           pop4_genes$gene),
  category = c("Pop 1",
               "Pop 2", 
               "Pop 3", 
               "Pop 4"),
  fill = c("#7d55a8",
           "#4d9817",
           "#ade187", 
           "#f1ec09"),
  height = 1000, 
  width = 1000,
  filename = NULL,
  main = "BLS"
)

grid::grid.draw(venn_p1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

pop_genes_intersect_2 <- intersect(
  intersect(intersect
            (pop1_genes$gene, pop2_genes$gene),
            pop3_genes$gene),
  pop4_genes$gene
  )


pop_genes_intersect_2
subset(bias_and_all, gene == pop_genes_intersect_2)

```

# pop1 pop3

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a Venn diagram 
library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(pop1_genes$gene, pop3_genes$gene),
  category = c("Pop1","Pop3"),
  fill = c("#187E93", "#C0D9AE"),
  height = 1000, width = 1000,
  filename = NULL,
  main = "BLS"
)

grid::grid.draw(venn_p1)
```


# All
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a Venn diagram 
#Colours:
# All samples #7d55a8
# 1 #187e93
# 2 #4d9817
# 3 #ade187
# 4 #f1ec09

library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(BLS_candidates$gene, pop1_genes$gene, pop2_genes$gene, 
           pop3_genes$gene,pop4_genes$gene),
  category = c("All samples","Pop 1", "Pop 2", "Pop 3", "Pop 4"),
  fill = c("#7d55a8", "#187e93", "#4d9817", "#ade187", "#f1ec09" ),
  height = 1000, width = 1000,
  filename = NULL,
  main = "BLS"
)

grid::grid.draw(venn_p1)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}

pop_genes_intersect <- intersect(
  intersect(
    intersect(
      intersect(
        BLS_candidates$gene, pop1_genes$gene),
      pop2_genes$gene),
    pop3_genes$gene),
  pop4_genes$gene
  )

subset(bias_and_all, gene == pop_genes_intersect)

```









--------------------------
  
3 Immune genes. 1 of them is in the supergene. 1 is a BUSCO and 1 is private. |
4 Supergene genes. 3 are BUSCO. 1 is immune. |
3 ASE. 2 are BUSCO. 1 is private. |
2 ORs. both private |

## ORS

### All ORs with the differen categories

```{r echo=FALSE,  message=FALSE, warning=FALSE}
candidate_immune <- subset(bias_and_all, ors == "1" & immune == "1")
candidate_busco <- subset(bias_and_all, ors == "1" & hymenop == "1")
candidate_super <- subset(bias_and_all, ors == "1" & super == "1")
candidate_ASE <- subset(bias_and_all, ors == "1"& ASE == "1")
candidate_ORs <- subset(bias_and_all, ors == "1") 
candidate_dros <- subset(bias_and_all, ors == "1"& candens != "0")

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a Venn diagram 
library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(candidate_immune$gene, candidate_busco$gene, candidate_super$gene, 
           candidate_ASE$gene, candidate_ORs$gene),
  category = c("Immune", "BUSCO", "Super", "ASE", "OR"),
  fill = c("#2A8DA1", "#C0D9AE", "#B7BAB2", "#C0D9AE", "#C0D9AE"),
  height = 1000, width = 1000,
  filename = NULL,
  main = "BLS"
)

grid::grid.draw(venn_p1)
```


### All ORs with the differen categories. Includind the BLS candidate ORs

```{r echo=FALSE,  message=FALSE, warning=FALSE}
candidate_immune <- subset(bias_and_all, ors == "1"& immune == "1")
candidate_busco <- subset(bias_and_all, ors == "1" & hymenop == "1")
candidate_super <- subset(bias_and_all, ors == "1" & super == "1")
candidate_bls_ors <- subset(bias_and_all, ors == "1" & candens != "0")
candidate_ORs <- subset(bias_and_all, ors == "1") 

candidate_bls_ors
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a Venn diagram 
library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(candidate_immune$gene, candidate_busco$gene, candidate_super$gene, 
           candidate_bls_ors$gene, candidate_ORs$gene),
  category = c("Immune", "BUSCO", "Super", "BLS_OR", "OR"),
  fill = c("#2A8DA1", "#C0D9AE", "#B7BAB2", "#C0D9AE", "#B45639"),
  height = 1000, width = 1000,
  filename = NULL,
  main = "Only ORs"
)

grid::grid.draw(venn_p1)


```


### All ORs with the differen categories. Includind the BLS candidate ORs

```{r echo=FALSE,  message=FALSE, warning=FALSE}
candidate_immune <- subset(bias_and_all, ors == "1" & immune == "1")
candidate_busco <- subset(bias_and_all, ors == "1" & hymenop == "1")
candidate_super <- subset(bias_and_all, ors == "1" & super == "1")
candidate_bls_ors <- subset(bias_and_all, ors == "1" & candens != "0")
candidate_ASE <- subset(bias_and_all, ors == "1" & ASE == "1")
candidate_ORs <- subset(bias_and_all, ors == "1") 


```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a Venn diagram 
library(VennDiagram)
venn_p1 <- venn.diagram(
  x = list(candidate_ASE$gene, candidate_busco$gene, candidate_super$gene, 
           candidate_bls_ors$gene, candidate_ORs$gene),
  category = c("ASE", "BUSCO", "Super", "BLS_OR", "OR"),
  fill = c("#2A8DA1", "#C0D9AE", "#B7BAB2", "#C0D9AE", "#B45639"),
  height = 1000, width = 1000,
  filename = NULL,
  main = "Only ORs"
)

grid::grid.draw(venn_p1)

```


```{bash echo=FALSE}
rm -rf Venn*log

```